---
title: Tidying Unkempt Data
author: Rhys Tilford
description: Read about how I used R to tidy and display urban transport data
date: 11/13/2023
categories: [Data Analysis, Programming, R]
---

```{r}
#| include: false
### load necessary packages
library(tidyverse)
library(rvest)
library(magrittr)
library(rlang)
library(stats)
```

### The Data

I found a fascinating data table on [this](https://en.wikipedia.org/wiki/List_of_metro_systems) Wikipedia page so I decided to pull it and analize it. Here is the code I used to pull it:

```{r}
#| warning: false
web_page <- "https://en.wikipedia.org/wiki/List_of_metro_systems"

Tables <-
  web_page %>%
  read_html() %>%
  html_elements("table")

MetroList <- 
  Tables %>% 
  purrr::pluck(1) %>%
  html_table()
```

Much to my dismay the data was riddled with miscellaneous junk. Here is the data inside MetroList:

```{r}
#| warning: false
#| echo: false
MetroList %>% 
  DT::datatable(class = "table table-hover",
                style = "bootstrap",
                options = list(pageLength = 5))
```

Here are the problems I see in MetroList:

 - The name of the column that holds the number of stations on each system has a long and irrelevent name.
 - Several of the columns have units and footnote links in them.

I decided to import it to Rstudio and take a crack at cleaning it.

### Tidying

Optimistically, I decided to call this new table TidyMetroList.

##### Renaming Columns

The following code renames the columns with a unified naming system:

```{r}
#| eval: false
MetroList %>% 
  set_names(c("City", "Country", "Name", "ServiceOpened", "LastExpanded", "StationCount", "Lines", "SystemLength", "RidershipMillions"))
```

```{r}
#| echo: false
MetroList %>% 
  set_names(c("City", "Country", "Name", "ServiceOpened", "LastExpanded", "StationCount", "Lines", "SystemLength", "RidershipMillions")) %>%
  DT::datatable(class = "table table-hover",
                style = "bootstrap",
                options = list(pageLength = 5))
```

##### 





